The brightness and spatial distributions of terrestrial radio sources

### 2013-01-07 articles/Article-RFI-distribution/article-rfi-distribution.tex

\begin{abstract}
Feeble undetected sources of radio-frequency interference (RFI) might become visible after long time integration of radio observations if they are coherently present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest objects in the Universe. We explore the low-frequency RFI situation by studying the $\log N$--$\log S$ histograms of data observed with LOFAR. The histograms shows an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law would imply that the current LOFAR EoR observations should be severely RFI limited if these sources are coherently present. Since these observations have achieved the thermal (sky) noise limit, we conclude that it is probable that most of the RFI sources do not add coherently.
\end{abstract}

>>>>

\begin{abstract}
Feeble undetected sources of radio-frequency interference (RFI) might become visible after integration of long radio observations if they are coherently present over time. Thereby, they might obstruct the detection of the astronomical signal of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals of the earliest objects in the Universe. We explore the low-frequency RFI situation by studying the $\log N$--$\log S$ histograms of data observed with LOFAR. The histograms shows an RFI distribution that follows a power-law distribution with an estimated exponent around -1.6 to -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law would imply that the current LOFAR EoR observations should be severely RFI limited if these sources are coherently present. These observations almost reach the thermal noise and are thought not to be limited by RFI. We conclude that it is likely that most RFI sources do not add coherently.
\end{abstract}
